---
title: "rola-r-tutorial"
output: html_document
date: "2025-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# STEP 0: Data Description

## Research Design

This simulated project represents a **clinical trial-style time-series study** analyzed under a **Difference-in-Differences (DiD) framework**. A total of 100 patients are observed daily for 100 days, with multiple electronic health record (EHR) check-ups per day recording heart rate. At the halfway point of the trial (day 50), an intervention is introduced: patients in the treatment group begin receiving a new drug intended to reduce elevated heart rate, while control patients remain untreated.  

- **Groups:** Patients are split into treated (receive drug at day 50) and control (never treated).  
- **Periods:** Pre-treatment (days 1–49) vs post-treatment (days 50–100).  
- **Treatment Effect:** The DiD estimator compares the change in average heart rate before vs after the intervention across treated and control patients. A greater reduction among treated patients relative to controls indicates the causal effect of the drug.  

This design mirrors real-world EHR studies where an intervention is rolled out at a fixed time, and differences in pre/post changes between groups are used to estimate impact.

---

## ehr_did.csv

This file contains the **full dataset** for analysis, combining demographics, trial structure, and repeated EHR measures into a single table with >10,000 rows. Each record corresponds to one check-up for a patient at a specific day and time. Key variables include:  

- **patient_id:** Unique identifier for each patient.  
- **treated:** Indicator (1 = treatment group, 0 = control group).  
- **day:** Day of follow-up (1–100).  
- **post:** Indicator (1 = post-intervention period, day ≥ 50; 0 = pre-intervention).  
- **datetime:** Exact timestamp of the check-up (to hour/minute).  
- **hr:** Heart rate (beats per minute) recorded at that check-up.  
- **sex, race, age, bmi, baseline_hr:** Demographic and baseline covariates carried into each row for convenience.  

This dataset can be aggregated to daily averages for visualization (e.g., mean daily HR by group) or analyzed as a full longitudinal panel for formal DiD estimation.


# R Pipeline

## STEP 1: LOAD PACKAGES

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(lme4)
library(rstanarm)
library(rdrobust)
```

## STEP 2: LOAD DATA

```{r}
ehr_did          <- read_csv("ehr_did.csv")

```

## STEP 2: Data Wrangling

```{r}
analysis_df <- ehr %>%
  mutate(date = as.Date(ymd_hms(datetime))) %>%
  group_by(patient_id, day, treated, post, date) %>%
  summarise(mean_hr = mean(hr), .groups = "drop")
```


## STEP 3: ANALYSIS

```{R}
fit <- lm(mean_hr ~ treated * post, data = analysis_df)
summary(fit)

plot_df <- analysis_df %>%
  group_by(day, treated, post) %>%
  summarise(avg_hr = mean(mean_hr), .groups = "drop")

ggplot(plot_df, aes(x = day, y = avg_hr, color = factor(treated))) +
  geom_line(size = 1.2) +
  geom_vline(xintercept = 50, linetype = "dashed", color = "red") +
  labs(
    x = "Day of Trial",
    y = "Heart Rate (bpm)",
    color = "Treated"
  ) +
  theme_bw()

```

## STEP 4: CONCLUSION
The Treatment is Highly Effective p < 0.05.


